
\section{Coherence}
\label{sec:coherence}

\subsection{The need of more coherence}
In the previous sections we showed how to define all coherence cells
witnessing the axioms of a strict $\omega$-groupoid. These can be
composed to witnesses identity of cells which don't exactly match the
sides of either of the axioms.

For instance, to witness the equality
\[(g \,\id_{b})\,f \quad = \quad g\, f\]
we can compose $\lambda$ after $\alpha$ to obtain witnesses such as: 
\[\id_{g}\lambda_f \cdot \alpha_{g,\id_g,f}  ~:~ (g
\,\id_{b})\,f \longrightarrow   g\, f\qquad\text{   and    }\qquad
 \rho \, \id_f ~ : ~ (g\,\id_{b})\,f \longrightarrow   g\, f\quad.\]

 In the strict case equality of 1-cells is a proposition and therefore
 all proofs of equality of 1-cells are equal. In the weak
 2-categorical case, equality of 1-cells is not propositional but
 equality of 2-cells is. In other words, equality of 1-cells is
 witnessed by 2-cells which must satisfy new axioms.  For instance: 
%
\begin{equation}\label{eq:lambda-alpha}
\bfig
\qtriangle<750,500>[(g\,\id_{b}) \,f ` g\,(\id_{b}\,f)`g\, f;
\alpha_{g,\id_g,f}`\rho\,\id_f`\id_g\,\lambda_f]
\efig
\end{equation}
% 
In our case, when all levels are weakened the equality
is replaced be a new 3-cell 
\[\id_{g}\lambda_f \cdot \alpha_{g,\id_g,f}
\longrightarrow \rho\,\id_f\]
%
Moreover, for any pair of such 3-cells, $p,\,q : \id_{g}\lambda_f \cdot \alpha_{g,\id_g,f}\longrightarrow
\rho\,\id_f$, there must be 4-cell $p \longrightarrow q$,  etc., all
the way up to $\omega$. This is a weakening of propositionality of
equality of $n$-cells in the strict setting. The following diagram
illustrates this up to level 4:
\begin{equation}\label{eq:lambda-alpha}
\bfig\scalefactor{1.5}
\qtriangle<750,500>[(g\,\id_{b}) \,f ` g\,(\id_{b}\,f)`g\, f;
\alpha_{g,\id_g,f}`\rho\,\id_f`\id_g\,\lambda_f]
\morphism(650,450)/{@{>}@/_1em/}/<-200,-200>[`;p]
\morphism(700,400)|b|/{@{>}@/^1em/}/<-200,-200>[`;q]
\morphism(480,400)|b|/..>/<170,-135>[`;]
\efig
\end{equation}
%
In summary and full generality: 
\begin{quote}
  \emph{For any pair of coherence cells with the same domain and
    codomain, there must be a mediating coherence cell.}
\end{quote}

\subsection{Formalising coherence cells between coherence cells}

%
We formalise the above principle as follows. We define a predicate 
$\thin$ on objects
% \[
% \AxiomC{$x :  \Obj~C$}
% \UnaryInfC{$\thin~x~:~\Set$}
% \DisplayProof
% \]
% %
such that $\thin\,(\id\,f)= \top$ where $\top$ is the one element Set. 
Moreover each coherence cell is thin
\[
\begin{array}{c}
\thin\,(\lambda\,\_\,\_) = \top \quad \thin(\rho\,\_\,\_)
  ~=~\top\quad \thin\,(\alpha,\_\,\_\,\_) = \top
  \quad\thin~(\chi\,\_\,\_\,\_\,\_) = \top\\
\thin\,(\iota~f) = \top \quad \thin\,(\kappa~f) = \top
\end{array}
\]
Further we close $\thin$ under weakening, composition and inverses. 
Finally, we introduce a constructor of $\Obj$ and a clause of $\thin$:
\[
\AxiomC{$f ~ g : \Obj~\homcat{C}{a}{b}$}
\AxiomC{$p : \thin~f$}
\AxiomC{$q : \thin~g$}
\TrinaryInfC{$\mathsf{coh}~p~q :
  \Obj~\homcat{\homcat{C}{a}{b}}{f}{g}$}
\DisplayProof
\qquad
\AxiomC{$\thin\,(\coh\,p\,q) = \top$}
\DisplayProof 
\]
The last remaining case are variables, which are not thin:
\[
\AxiomC{$\thin\,(\var\, v) = \bot$}
\DisplayProof 
\]

\subsection{The problem with coherence}
The question of coherence in weak $\omega$-categories is subtle. On
the one hand, one needs enough coherence to make weakly equivalent all
cells that should be identities in the strict case. On the other
hand, if we shouldn't add too many equations not to loose any of our
intended models. We believe our definition is correct as we are saying in
the definition of $\mathsf{coh}$ only that all \emph{formal} diagrams
of coherence cells commute (see \cite{maclane}, \S VII.2, for a
related discussion). The fact that not all diagrams commute rests on
the fact that one is not allowed to postulate equations between
nonvariables, which is achieved by the separation of $\VarCat$ and
$\Cat$ and the fact that contexts are built from $\VarCat$s. Otherwise
it would be for instance possible\footnote{We are indebted to an
  anonymous referee for pointing out this example to us.}  to assume a
0-cell $a$, and a pair of 2-cells: $x, y : \id~a \rightarrow \id~a : a
\rightarrow a$. Then it is be possible to construct, by the Eckerman-Hilton
argument, a cell $\eta : x \circ y \rightarrow y \circ x : \id~a
\rightarrow \id~a : a \rightarrow a$ which is thin. As $\id~(x\circ
y)$ is also thin, $\eta\circ\eta$ and $\id (x \circ y)$ would be
equivalent by $\coh$. However, not every monoidal braided category is
symmetrical. However  it is not possible to construct this example in our
syntactical framework as $(\id~a)$ is not a variable.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "weakomega-csl"
%%% End: 
