\section{Conclusions and Further Work}
\label{sec:conclusions}

\subsection{Summary}
We have presented a novel approach to defining weak $\omega$-groupoids
which is based on ideas from Type Theory. The central idea is to
define the syntax of weak $\omega$-groupoids and then define a weak
$\omega$-groupoid as a globular set with an interpretation of the
syntax, which is where Type Theory has its greatest strength. Our
approach to formalization of coherence is natural, in a way naive,
since it is a natural generalisation of the corresponding first order
laws.

We have formalized all of the material presented here in Agda
\cite{agda}, except for the definition of the telescope morphism for
$\chi$.
% , which wasn't ready in time for submission.     
We believe this is
a technicality, albeit a difficult one. The Agda source file is available
from \cite{coreagda}.


\subsection{Related work}
There exists an abundance of categorical definitions of weak
$\omega$-categories and groupoids.  Although a direct comparison is a
slippery road, we would like to give a rough comparison of the key
similarities and differences between our and other definitions. Most
importantly, our definition is fundamentally different in that it is
formulated in Type Theory rather than Category Theory. This means that
we couldn't just formalise any of the approaches
\cite{penon:1999,batanin98:monoidal-globular,leinster:2000} because
the notion a strict $\omega$-category is central in them in that it
drives the definition of coherence cells. However, it is unclear to how define
\emph{strict} $\omega$-categories in Type Theory
without quotient types. This forces some of the choices we have
made. Nevertheless, on an intuitive level there are similarities of
our approach to some 
categorical approaches, in particular to Batanin's definition
\cite{batanin98:monoidal-globular} which we briefly discuss below:  
\begin{itemize}
\item Batanin's \emph{spans} are essentially our telescopes. And as
  noted by Batanin, Cartier called Batanin's spans telescopes in
  1994. 
\item Batanin's definition, same as ours, is globular and works with a system of
  units and binary compositions. 
\item We conjecture that for a $C : \Cat~\Gamma$, the mapping 
  $n \mapsto \Sigma(t : \Tel~C~n) (\Obj (t \Downarrow))$
  forms a strict monoidal globular category freely generated by the
  globular set determined by $\Gamma$. 
\item It remains to show that our syntax defines a contractible
  $\omega$-operad and our notion of interpretation defines its
  algebra.
\end{itemize}
%



\subsection{Further work}
The current formalisation is still quite complicated and we hope to
find ways to simplify it. One interesting idea may be to use the
syntactical approach to define opetopes based on dependent polynomial
functors (i.e. indexed containers) \cite{opetopes}, which has a very
type-theoretic flavour.

It remains to prove that the definition proposed in this paper is a
sensible one. This seems to be most easily done by showing that any
interpretation of the syntax defines an algebra for Batanin's weakly
initial contractible operad. 

We would like to use our framework to provide a formalisation of a
variation of the results in
\cite{lumsdaine10:weak-o-categories,berg08:types-are} by showing that
$\Idw$ is a weak $\omega$-groupoid. Such a formalisation would be
different form their results because we are working inside Type
Theory, rather than on a meta-level.

The main challenge ahead is to formalize the notion of a
$\omega$-groupoid model of Type Theory. Once this has been done we
will be able to eliminate the univalence axiom and provide a
computational interpretation of this principle.





%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "weakomega-csl"
%%% End: 
